Generating presidential tweets with an LSTM network

### 2018-04-22 paper/paper.tex
Add finished paper.
I've never hated a paper before. This is a new feeling.

\begin{abstract}
    \todoinline{Write an abstract, but after the paper is finished.}
\end{abstract}

@@

@@

        \begin{figure}[h]
            \centering
            \begin{tikzpicture}[scale=0.25]

@@

        \newpage
        % I have a new hatred for floating wrapped figures. Welcome to a new kind of hell.
        \begin{wrapfigure}[50]{r}[20pt]{0pt}
            \centering
            \includegraphics[width=0.39\textwidth]{figures/64x6.eps}
            \caption{A deep and narrow LSTM model using batch normalization}\label{fig:archetecture-64x6}
        \end{wrapfigure}

@@

        Also note that the author trained the experimental models on \textit{all} of President Trump's tweets from his account creation in 2009 to 2018.\todo{Move this to discussion}{}

    \subsection{Model Architectures}
        The author attempted 21 different variations of LSTM models in both height and depth. Most examples and online recommendations were to use relatively shallow networks with height determined by circumstance.\todo{Citation needed?}{}

        One such configuration that showed promise is shown in~\autoref{fig:archetecture-64x6}. This model was composed of six 64-unit LSTM layers with batch normalization. Each configuration used 60 character long sequences as the model input, and a one-hot encoded vector output to indicate the most likely next character in the sequence.

@@

        When using RMSProp, it was typically only necessary to train for around 5 epochs, depending on the model size, while SGD (even with a higher learning rate) took on average 100 epochs to get comparable results.

@@

\section{Results}
    \todoinline{
        \begin{itemize}
            \item Hack validation data into training and produce plots?
            \item Give example tweets?
            \item Grammar, spelling, etc.
            \item Accuracy is meaningless because you're trying to create \textit{new} tweets. This is where the sampling with diversities comes in. The higher diversity the more creative the network is, but the more meaningless it becomes.
            \item Produce a largeish amount of generated tweets.
        \end{itemize}
    }
    \subsection{256 Unit, 1 Layer LSTM}

        \begin{figure}[h]
            \centering
            \includegraphics[width=0.4\textwidth]{figures/256dropout.eps}
            \caption{256 unit LSTM model with dropout}\label{fig:archetecture-256}
        \end{figure}

    \subsection{64 Unit, 6 Layer LSTM}
        The following results are using the model shown in~\autoref{fig:archetecture-64x6}.

    \subsection{1024 Unit, 1 Layer LSTM}
    \begin{figure}[h]
        \centering
        \includegraphics[width=0.5\textwidth]{figures/1024batchnorm.eps}
        \caption{1024 unit LSTM model with batch normalization}\label{fig:archetecture-256}
    \end{figure}

    \subsection{Textgenrnn}
    \begin{figure}[h]
        \centering
        \includegraphics[width=\textwidth]{figures/context_model.png}
        \caption{Textgenrnn architecture}\label{fig:archetecture-256}
    \end{figure}

\section{Discussion}
    \todoinline{
        \begin{itemize}
            \item How good was it?
            \item Accuracy meaningless because attempt is creative generation.
            \item Did it pick up on patterns backed by data?
        \end{itemize}
    }

\section{Future Work}
    \todoinline{
        \begin{itemize}
            \item Word level generation. Instead of each character being mapped to an integer, create a vocab list and map each word to an integer.
            \item Word association norms?
            \item Embeddings?
            \item Combining sources?
            \item Grammar and spelling, part of speech tagging, word association, linguistic features as a component of a GAN
            \item Crowd-sourced adversarial network as a training component?
        \end{itemize}
    }

>>>>

\begin{abstract}
    We motivate and develop a recurrent generational model to generate tweets similar to those of President Donald J. Trump. We outline previous research in generational models, and point out the difficulties of objective measurement of inherently subjective creative models. After experimenting with several recurrent models, we settle on a 6 layer, 64 unit LSTM model, and recommend a more advanced model as proposed by other research. We finally recommend using advanced natural language processing techniques to develop a Generative Adversarial Network in any future work.
\end{abstract}

@@

        \begin{figure}[H]
            \centering
            \begin{tikzpicture}[scale=0.25]

@@

        % I have a new hatred for floating wrapped figures. Welcome to a new kind of hell.
        % Also, tall and skinny figures just aren't a good idea...
        \begin{wrapfigure}[47]{r}[40pt]{0pt}
            \centering
            \includegraphics[width=0.36\textwidth]{figures/64x6.eps}
            \caption{A deep and narrow LSTM model using batch normalization}\label{fig:architecture-64x6}
        \end{wrapfigure}

@@

@@

        Also note that the author trained the experimental models on \textit{all} of President Trump's tweets from his account creation in 2009 to 2018.

    \subsection{Model Architectures}
        The author attempted 21 different variations of LSTM models in both height and depth. Most examples and online recommendations were to use relatively shallow networks with height determined by circumstance.

        One such configuration that showed promise is shown in~\autoref{fig:architecture-64x6}. This model was composed of six 64-unit LSTM layers with batch normalization. Each configuration used 60 character long sequences as the model input, and a one-hot encoded vector output to indicate the most likely next character in the sequence.

@@

        When using RMSProp, it was typically only necessary to train for around 5 epochs, depending on the model size, while SGD (even with a higher learning rate) took on average 100 epochs to get comparable results. All attempts used a categorical cross-entropy loss function due to the network output being one-hot encoded.

@@

\section{Results}
    The results for a few select architectures are summarized below.

    \subsection{256 Unit, 1 Layer LSTM}
        \autoref{fig:256-loss} shows the training loss over time for a basic one-layer 256 unit LSTM model. The author used validation data to get a sense of how well the models generalized. The author did \textit{not} use a test set to evaluate how the model performed at the end of training because of the problem nature. Generational models are not intended to accurately map the training inputs to the training outputs. Thus evaluating the model accuracy on an unseen dataset gives no useful information.

        \begin{figure}[h]
            \centering
            \includegraphics[width=0.7\textwidth]{figures/256dropout20.png}
            \caption{Basic 256 unit LSTM model}\label{fig:256-loss}
        \end{figure}

        Instead, the author used subjective measures to evaluate whether a model performed well or not. The output of this network, for lower sampling diversities, was quite intelligible. Several generated tweets have been cherry-picked, as \citet{karpathy_2015} notes is necessary for recurrent generational models, and shown below.

        Note that all capitalization has been done by hand to make the results more legible.

        \begin{quote}
            ``The debate to the worst and the worst President Obama was a great president of the world is a great person. The Republicans will be a great job on the world''

            ``The state of the state of the world is a great person. The Republicans will be on the world is a great poll to be the best interview discussing the Trump Tower''

            ``@realDonaldTrump is the only one of the worst president of the world is a great person. The Republicans will be a great president of the world''

            ``I have a great people of the worst and the debate to the worst President Obama is a great person. The Democrats are all talk about me to be the best interview discussing the Trump Tower''

            ``The poor for their presidential party of his presidency to start doing for millions of the conservative and brand who has got a state of the good country! We need a state who says Hillary she are an amazing''
        \end{quote}

        It is quite evident that the network is catching on to certain speech patterns, but is completely blind to grammar, and only strings together cogent thoughts over small stretches of the generated text. When considering the sentence structure as a whole, it lacks meaning, but certain substrings (e.g., ``the worst President Obama was a great president'', or ``The Democrats are all talk about me'') form coherent thoughts.

    \subsection{64 Unit, 6 Layer LSTM}
        The following results are using the model shown in \autoref{fig:architecture-64x6}. Note that \autoref{fig:64x6-loss} shows that this model took more training time to reach the same behavior as shown in \autoref{fig:256-loss}. This is expected because it's a larger model.

        \begin{figure}[h]
            \centering
            \includegraphics[width=0.7\textwidth]{figures/64x6-rms-batchnorm-50.png}
            \caption{64 unit, 6 layer LSTM model with batch normalization}\label{fig:64x6-loss}
        \end{figure}

        Again, several cherry-picked tweets are given below with capitalization added.

        \begin{quote}
            ``@MittRomney should be the only one who need it. Good move for the Scott exciting to Russia on @FoxNews for the administration policies commissioner''

            ``Dems is a total loser with the Trump country is a great guy to the great states of The Apprentice''

            ``@BarackObama scandaled will be Donald Trump liars will be a game on @foxandfriends at Americans and winner!''

            ``Bad interview to the Trump country is the states of the first story and states of the Trump International Hotel in the country will be interviewed on @FoxNews and never win in the presidential''

            ``Statements are something to be a truly great guy to the only one who can speak at the Trump Tower to see the president of the US''

            ``Trump country is a great president of the USA and the art of the Democrats are something to be a great support the Trump Tower to make America great again!''

            ``The end of Monday at your cards classing a billionaire inspiration''

            ``And the Trump country is a great president of the USA and the Trump signing to stop president of the USA and the Trump signing to stop and the Trump signitictor is a great president of the USA and states and states of the Trump country is a great president of the USA''
        \end{quote}

    \subsection{1024 Unit, 1 Layer LSTM}
        During the experimentation process, it seemed like the shallow layers produced the most intelligible results, so the author experimented by making the models much much wider. However, the results were not promising. \autoref{fig:1024-loss} shows the training loss over time, which shows that the optimizer struggled with such a large model. Further, training even for 20 epochs took almost 24 hours on an NVIDA GTX 1080.

        \begin{figure}[h]
            \centering
            \includegraphics[width=0.7\textwidth]{figures/1024-rms-batchnorm-20.png}
            \caption{1024 unit LSTM model with batch normalization}\label{fig:1024-loss}
        \end{figure}

        The generated results are given below.

        \begin{quote}
            ``and and and and and and and the and and and and and the and and the and and and and and and and and and and and and and the and and and the and the and the and and the and and the and and and the and and and and and and the and the and than the and and the and and and and and''

            ``@realDonaldTrump i we aa seare to need a fight and and any every is a tered beanies and the of the real in the life and that I live a great and leader I love to the beal about is any and and the me. The thanes and that I and and and lime''

            ``America get I and the thang is and the ane that I at see any I and than Obama in is and peolle last is than the antary this peesident a. The unition so to thanised is a taine in the really and it in I've any way a deal to the for the aldonal and was and we will be a like''
        \end{quote}

    \subsection{Textgenrnn}
        After it was too late to incorporate any meaningful changes, the author was directed to a GitHub project \citep{woolf_2017} built expressly for generating text using a recurrent network. The model architecture is shown in \autoref{fig:architecture-textgenrnn}. This architecture is interesting because it has a context input to allow mixing tagging different pieces of the training set with context. The example usage given in the documentation indicates that this works well when combining text written by different authors.

        \begin{figure}[h]
            \centering
            \includegraphics[width=\textwidth]{figures/context_model.png}
            \caption{Textgenrnn architecture}\label{fig:architecture-textgenrnn}
        \end{figure}

        For completeness, Textgenrnn, as trained by the author, may be used as follows

        \begin{minted}{python}
            from textgenrnn.textgenrnn import textgenrnn  # Installable with Pip
            # Files found in trumperator/experiments/
            t = textgenrnn(config_path='trumperator_config.json',
                           vocab_path='trumperator_vocab.json',
                           weights_path='trumperator_weights.hdf5')
            t.generate(1, temperature=0.2)
        \end{minted}

        Textgenrnn has the option of tokenizing on either words or characters. The following generated samples were created using character-tokenized sequences. Note that any capitalization is \textit{not} added by the author---it was generated by the model. Also note that the documentation stated to train the model for a mere five epochs.

        \begin{quote}
            ``@realDonaldTrump you would be a clue and it's a great season. The Second and a state and all why are you all there, a real estate of him right on his interests and entertaining the U.S. and the world and the great policies are a fantastic game!''

            ``I will be on @foxandfriends at: A.M. Enjoy!''

            ``The Trade Resort is killing our country picks at the borders that we can energy the president of a total leader and a winner of the most broke our since I all television clifful in the other politics of the borders are a great complete disaster.''

            ``@realDonaldTrump the many of the Republican candidates and have been a big problem. \#Trump''

            ``@realDonaldTrump @FoxNews  You are the only one that are with your badly because you support it. The man at our country will be great to have the mother is one of the big millions of comments are looking like a big people of the U.S. he will be so much more than he is the U.S.''

            ``Highly improvement on back with Democrats of the coming evenings than the lands showication and one of the place tomorrow at: A.M.  Enjoy!''
        \end{quote}

        These results seem to be more coherent, but it does seem like it simply strings together substrings from the larger corpus.

\section{Discussion}
    Even taking the entirety of President Trump's tweets (including time not spent in office), cherry-picked results have a quite political flavor. This was unexpected, but welcome as it exacerbated the patterns shown in the generated corpus. Even so, it is difficult to get a good objective grasp of how well generational models perform due to the qualitative nature of their outputs. Therefore, accuracy analysis is essentially meaningless, especially because the intent of generational models is to \textit{avoid} exactly replicating instances in the training set.

    We ultimately recommend using the six-layer, 64 unit LSTM network illustrated in \autoref{fig:architecture-64x6} because the generated output seemed to be more diverse, with less repetition. However, we do have recommendations for future work in this area. There has been an enormous amount of recent research in natural language understanding using machine learning models. The author suggest any further research look into using spaCy \citep{spacy2} to strip each word to its root form, perform part-of-speech tagging, grammar analysis, word association analysis, and other linguistic analysis as a one or more components of a Generative Adversarial Network (GAN).

    The best performing models as a whole generated fairly reasonable results, but as \citet{karpathy_2015} notes, most recurrent generational models produce results that must be scraped by hand to find the best results. We recommend a different strategy for future work however. It seems like the Textgenrnn model could be used as the generative portion of a Generative Adversarial Network, and use some kind of grammar metric as the adversarial component of the model.

    Further, the author also suggests tokenizing at a word level so that the recurrent model predicts the next word in a sequence rather than the next character. This is drastically increase the size of the output vector, but will diminish the size of dataset. It will probably be necessary to prune out the least common words, as well as treat punctuation as individual tokens.

    It would also be possible, if training time is not an issue, to crowd-source the adversarial component of a GAN by tweeting each generated tweet, and recording how many likes or retweets each received.
